#import "@local/bootlin:0.1.0": * 

#import "@local/bootlin-yocto:0.1.0": *

#import "@local/bootlin-utils:0.1.0": *

#import "../../typst/local/themeBootlin.typ": *

#import "../../typst/local/common.typ": *

#show: bootlin-theme.with(
  aspect-ratio: "16-9",
config-common(
  handout: "handout" in sys.inputs and sys.inputs.handout == "1",
))

#show raw.where(block: true): set block(fill: luma(240), inset: 1em, radius:0.5em, width:100\%) 

#show raw.where(block: false): r => { text(fill: color-link)[#r] } 

 == alsa-utils

=== {alsa-utils}
  \begin{itemize}
  \item ``` alsa-utils ``` is a repository of tools to interact with
    ALSA devices
  \item \url{https://github.com/alsa-project/alsa-utils}
  \end{itemize}


=== {Controls}
  \begin{itemize}
  \item ``` alsamixer ``` provides a ``` ncurse ``` based graphical
    interface to modify sound cards controls.
  \item ``` amixer ``` is a command line tool to set controls.
    \begin{itemize}
    \item The ``` scontrols ``` and ``` controls ``` commands list the
      available controls.
    \item The ``` contents ``` commands list the
      available controls and shows their content.
    \item The ``` cset ``` and ``` sset ``` commands allows modifying
      the controls.
    \item The ``` cget ``` and ``` sget ``` commands show the content of
      a specific control.
    \end{itemize}
  \item ``` alsactl ``` is a tool that can save the control values to a
    file and restore them from a file.
  \end{itemize}


=== {Playback and capture}
  \begin{itemize}
  \item ``` speaker-test ``` can generate tones or noises to play on
    specific channels with a specified rate.
  \item ``` aplay ``` plays an audio file. It is able to set many
    stream parameters.
  \item ``` arecord ``` can record an audio stream to a file.
  \end{itemize}


#setupdemoframe([Userspace tools],[
  Using userspace tools to:
  \begin{itemize}
  \item configure sound card controls
  \item load and store default values for controls
  \item play sound
  \item record
  \end{itemize}
])